Atmospheric escape and star-planet-interaction in the HAT-P-67 b system

### 2022-11-28 paper/paper1/ms.tex
writing the analysis

The Habitable Zone Planet Finder Spectrograph \citep[HPF][]{2012SPIE.8446E..1SM,2014SPIE.9147E..1GM, 2019Optic...6..233M} on the queue-scheduled 10-meter \emph{Hobby-Eberly Telescope} \citep[HET][]{1998SPIE.3352...34R} operates in the near-IR from $8100-12800~\AA\;$ spanning the \textit{z}, \textit{Y}, and \textit{J} bands at spectral resolving power $R=55,000$. The HET fixed-elevation design \citep{2007PASP..119..556S} limits the observability of HAT-P-67 to less than 1 hour ``tracks'' for a fixed range of hour angles before (east track) and after (west track) transiting the meridian.  Whereas conventional steerable telescopes could conduct cintinuous point-and-stare observations of HAT-P-67 for hours, HET cannot.  In practice, this limitation means that in-transit and out-of-transit were rarely possible on the same night.  Instead, we organized the observations into four campaigns to coincide with HAT-P-67b transits on the nights beginning 2020 April 28, 2020 May 22, 2020 June 15, and 2022 April 28.  These in-transit nights had up to 14 exposures per HET track, with integration times between 5-8.5 minutes.  We obtained out-of-transit baseline observations up to 2 nights before and after the transit, plus random-in-phase low priority P4 queue-filler observations.  These out-of-transit observations typically received 4 or fewer individual exposures.  We observed HAT-P-67 with HPF for a total 41 visits on 39 unique nights, with two of those nights observing both an east and a west track.  Table \ref{tabHPFLog} summarizes the log of observations.

@@

We masked spectral regions predicted to have significant telluric lines with a template generated by \texttt{telfit} \citep{2014AJ....148...53G}.  We shifted the spectral coordinates to their common barycentric-corrected reference frame \citep{2014PASP..126..838W} as implemented in \texttt{astropy} \citep{2013A&A...558A..33A,2018AJ....156..123A}.  The sky-subtraction, telluric masking, barycentric correction, and all of our other standard pre-processing steps are implemented in the open-source Python interface \texttt{muler} \citep{2022JOSS....7.4302G}.

Figure \ref{fig:HPFheliumOverview} shows a zoom-in on the He 10830 region of interest, with all 153 individual exposures overlaid.  We see a variability in the Helium line of up to 10\%, much greater than the $<1\%$ pixel-to-pixel variation.  The feature width spans about 3 $\AA~$, too broad to originate from the much narrower telluric lines.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT-P-67_2022_clean.png}
    \caption{HPF spectra of HAT-P-67 from the April/May 2022 campaign.  We see conspicuous absorption during and immediately preceding the transit, but not immediately after.}
    \label{fig:HPF2022campaign}
\end{figure}

@@

\subsection{Calcium IR Triplet lines}

>>>>

The Habitable Zone Planet Finder Spectrograph \citep[HPF][]{2012SPIE.8446E..1SM,2014SPIE.9147E..1GM, 2019Optic...6..233M} on the queue-scheduled 10-meter \emph{Hobby-Eberly Telescope} \citep[HET][]{1998SPIE.3352...34R} operates in the near-IR from $8100-12800~\AA\;$ spanning the \textit{z}, \textit{Y}, and \textit{J} bands at spectral resolving power $R=55,000$. The HET fixed-elevation design \citep{2007PASP..119..556S} limits the observability of HAT-P-67 to less than 1 hour ``tracks'' for a fixed range of hour angles before (east track) and after (west track) transiting the meridian.  Whereas conventional steerable telescopes could conduct cintinuous point-and-stare observations of HAT-P-67 for hours, HET cannot.  In practice, this limitation means that in-transit and out-of-transit were rarely possible on the same night.  Instead, we organized the observations into four campaigns to coincide with HAT-P-67b transits on the nights beginning 2020 April 28, 2020 May 22, 2020 June 15, and 2022 April 28.  These in-transit nights had up to 14 exposures per HET track, with integration times between 5-8.5 minutes.  We obtained out-of-transit baseline observations up to 2 nights before and after the transit, plus random-in-phase low priority P4 queue-filler observations.  These out-of-transit observations typically received 4 or fewer individual exposures.  We observed HAT-P-67 with HPF for a total 41 visits on 39 unique nights, with two of those nights observing both an east and a west track.  The total on-source integration time exceeds 13.8 hours.  Table \ref{tabHPFLog} summarizes the log of observations.

@@

We masked spectral regions predicted to have significant telluric lines with a template generated by \texttt{telfit} \citep{2014AJ....148...53G}.  We shifted the spectral coordinates to their common barycentric-corrected reference frame \citep{2014PASP..126..838W} as implemented in \texttt{astropy} \citep{2013A&A...558A..33A,2018AJ....156..123A}.  We flattened the spectra to two pre-selected continuum indices highlighted as vertical blue bands in Figure \ref{fig:HPFheliumOverview}.  The sky-subtraction, telluric masking, barycentric correction, flattening, and all of our other standard pre-processing steps were carried out in the open-source Python interface \texttt{muler} \citep{2022JOSS....7.4302G}.

Figure \ref{fig:HPFheliumOverview} shows a zoom-in on the He 10830 region of interest, with all 153 individual exposures overlaid.  We see a variability in the Helium line of up to 10\%, much greater than the $<1\%$ pixel-to-pixel variation.  The feature width spans about 3 $\AA~$, comparable to the stellar rotationally lines too broad to originate from the much narrower telluric lines.

@@

The 13.8 hours of exposure, combined with HAT-P-67b's short ($P\sim4.8$ day) period, means that a large fraction of the orbit has been collected, with some phases (especially near transit) heavily sampled, and some other out-of-transit phases sampled only sparsely. We constructed a 2-D intensity phase scan, binning in phase and wavelength, and spanning the entire orbit of HAT-P-67.  Figure \ref{fig:HPFphase2D} overlays the planet's approximately $\pm150\;$km/s orbital Doppler velocity, with the stellar rest frame velocity demarcated by the vertical line.  The vanishing $<36$ m/s reflex motion of the star is imperceptible at this scale.  The horizontal dashed lines indicate the transit ingress, midcenter, and egress.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/phase_2D_diagram.png}
    \caption{Phase scan of intensity for HAT-P-67.  The spectrum flux intensity shows significant in-transit absorption during and before the transit.  Telluric lines redward of the line center have been masked.  An unmasked telluric line near 10825 $\AA$ can be seen.}
    \label{fig:HPFphase2D}
\end{figure}

A large absorption signal can be seen preceding transit ingress, with some significant absorption evident before $-0.2$ in phase.  We construct a fractional residual absorption spectrum by subtracting off and re-normalizing to the non-varying baseline.  We define the non-varying baseline spectrum as the average over phases $0.3-0.5$, which exhibited stable spectra with the lease absorption.  Figure \ref{fig:HPFscanResid} shows a conspicuous view of the Helium absorption of HAT-P-67b.  Significant absorption can be seen as a few percent at $-0.37$ in phase, rising to 10\% just before and during transit.  The egress drop off is extremely sharp, with almost no absorption directly after transit.

The absorption resides mostly near the stellar zero restframe velocity reference.  As mentioned previously, the telluric masking near 10836 $\AA$ censors our view of the existence (or not) of any redshifted lobe in this region.  A more careful treatment of the telluric absorption could rescue some of the data in this region.

Finally, a tentative much weaker post-egress bifurcation into blue-shifted absorption and redshifted emission may exist near 0.1 in phase.  The signal appears at or just above the noise level, but is shared among tens of pixels.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/phase_2D_diagram_resid.png}
    \caption{Absorption depth phase scan over the entire HAT-P-67b orbit from 41 visits with HPF.  }
    \label{fig:HPFscanResid}
\end{figure}

%\begin{figure}
%    \includegraphics[width=\linewidth]{figures/HAT-P-67_2022_clean.png}
%    \caption{HPF spectra of HAT-P-67 from the April/May 2022 campaign.  We see conspicuous absorption during and immediately preceding the transit, but not immediately after.}
%    \label{fig:HPF2022campaign}
%\end{figure}

@@

\subsection{Calcium IR Triplet and other line diagnostics}
We examined the Calcium IR Triplet lines (Ca IRT) for evidence of variability.  No variation was seen in these features, which were pre-processed in the same way as the Helium feature.  Figure \ref{x} shows a similar phase plot as Figure \ref{fig:HPFscanResid}, adapted to the line at
